---
title: "MSOCclassifier: an R package for multi-omics subtyping of high-grade serous ovarian cancer"
output: 
  BiocStyle::html_document:
    toc: FALSE
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{MSOCclassifier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
date: ""  # 去掉日期
---


<style>
h1.title {
  text-align: center;
}
</style>

<!-- <div style="text-align: center;"> </div> -->
Lingli He<sup>1</sup> and Xin Wang<sup>1</sup>

<sup>1</sup> Department of Surgery, The Chinese University of Hong Kong, Hong Kong SAR, China.

* contact: xinwang@cuhk.edu.hk
* `r paste("date:", Sys.Date())`
* package: `r BiocStyle::pkg_ver("MSOCclassifier")`



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# Introduction

The vignette helps the user to do multi-omics high-grade serous ovarian cancer subtyping using sparse mCCA (@witten2009extensions) and weighted average. Paired mRNA expression, microRNA expression, DNA methylation, copy number variation, and mutation data from TCGA-OV dataset were used for the training of multi-omics high-grade serous ovarian cancer classifier. The package accepts any combination of mRNA expression, microRNA expression, DNA methylation, copy number variation, and mutation data as input. 


# Package installation

Please run all analyses in this vignette under version 2.10 of R prior to installation of package MSOCclassifier, R packages caret should be installed. The package can be installed directly from CRAN (Comprehensive R Archive Network):
```{r message=FALSE, warning=FALSE, results='hide'}
options(repos = c(CRAN = "https://cloud.r-project.org/"))
install.packages("caret")
library(caret)
library(devtools)
# install the "MSOCclassifier" package
install_github("Carpentierbio/MSOCclassifier")
```

# A quick start 

## Projecting each omics data into a unified space and integrating them
The example dataset used in this analysis comes from the ICGC-OV cohort on 79 ovarian cancer patients, downloaded from https://dcc.icgc.org/projects/OV-AU (This link may no longer be accessible as of now).

* input data can be any combination of mRNA expression, microRNA expression, DNA methylation, copy number variation, and mutation data.
* if the single-omics or multi-omics integrated data includes mRNA expression or miRNA expression, *log*~2~\ transformation is required before integration.
* mRNA expression, microRNA expression, DNA methylation input data should be pre-processed and z-score normalized.

```{r, eval=TRUE}
options(knitr.duplicate.label = "allow")
library(MSOCclassifier)
library(dplyr)
# Load example multi-omics expression profile
data("rna_log_tpm_ICGC")
data("mir_ICGC")
data("methy_M_ICGC")
data("cnv_ICGC")
data("mut_ICGC")
# Load projection matrices derived from 226 TCGA-OV samples
data("TCGA_projection_mx")
# Load pre-processed TCGA-OV multi-omics data for feature selection in validation cohort
data("TCGAmRNAscaled")
data("TCGAmiRNAscaled")
data("TCGAmetscaled")
data("TCGAcnvscaled")
data("TCGAmutscaled")
# ensure that the number and order of features in the test data are identical to those in the training data (TCGA)
geneexp_bf_mapping = rna_log_tpm_ICGC[colnames(TCGAmRNAscaled), ]
geneexp_bf_mapping = t(scale(t(geneexp_bf_mapping)))
mirexp_bf_mapping = mir_ICGC[colnames(TCGAmiRNAscaled), ]
mirexp_bf_mapping = t(scale(t(mirexp_bf_mapping)))
methy_bf_mapping = methy_M_ICGC[colnames(TCGAmetscaled), ]
methy_bf_mapping = t(scale(t(methy_bf_mapping)))
cnv_bf_mapping = cnv_ICGC[colnames(TCGAcnvscaled), ]
cnv_bf_mapping = t(scale(t(cnv_bf_mapping)))
mut_bf_mapping = mut_ICGC[colnames(TCGAmutscaled), ]
mut_bf_mapping = t(scale(t(mut_bf_mapping)))
mut_bf_mapping[is.na(mut_bf_mapping)]=0
# Projecting each omics data into an unified space
mRNAexprCCA = t(geneexp_bf_mapping) %*% TCGA_projection_mx$ws[[1]]
mRNAexprCCA_2 = scale(mRNAexprCCA)
miRNAexprCCA = t(mirexp_bf_mapping) %*% TCGA_projection_mx$ws[[2]]
miRNAexprCCA_2 = scale(miRNAexprCCA)
methyexprCCA = t(methy_bf_mapping) %*% TCGA_projection_mx$ws[[3]]
methyexprCCA_2 = scale(methyexprCCA)
cnvexprCCA = t(cnv_bf_mapping) %*% TCGA_projection_mx$ws[[4]]
cnvexprCCA_2 = scale(cnvexprCCA)
mutexprCCA = t(mut_bf_mapping) %*% TCGA_projection_mx$ws[[5]]
mutexprCCA_2 = scale(mutexprCCA) # samples in rows and genes in columns 
# Multi-omics data fusion
a1 = a2 = a3 = a4 = a5 = 0.2
data_input = a1*mRNAexprCCA_2 + a2*miRNAexprCCA_2 + a3*methyexprCCA_2 + a4*cnvexprCCA_2 + a5*mutexprCCA_2
colnames(data_input) = paste("X",1:ncol(data_input),sep = "")
 
```

## Multi-omics high-grade serous ovarian cancer subtype classification
The `classifyMSOC` function requires an expression matrix with samples in rows and multi-omics features in columns. The column names of the expression profile should be X1, X2, ..., X100. The code chunk below demonstrates how to perform classification using primary high-grade serous ovarian cancer example data. 

```{r, eval=TRUE, message=TRUE}

# MSOC prediction of primary high-grade serous ovarian cancer
result <- classifyMSOC(data_input)
label <- result$label
prob <- result$prob %>%
  `colnames<-`(paste("MSOC", 1:5, "_prob", sep = ""))
res <- data.frame(prob, subtype = paste("MSOC", label, sep = "") ) %>%
  `rownames<-`(names(label))
head(res)
```


# Session Info
```{r, echo=FALSE}
  sessionInfo()
```

# References
